---
title: "ss"
output: html_document
date: "2023-10-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fitting an Autoregressive model to US quarterly GDP data

```{r}
# Load libraries
library(haven)
library(fpp3)
library(zoo)
library(patchwork)
library(tidyquant) 
```

```{r}
# Set parameters for graphical view
par(mfrow=c(2,1))
```

```{r}
# Load data
US_QGDP <- read_dta("C:/Users/idkup/Desktop/Course studies/2022 TERM 2/ECON4120/US_QGDP.dta")
head(US_QGDP)
```

```{r}
# Check for NA
sum(is.na(US_QGDP))
```

### Converting to tsibble object

Since the original dataset was created for STATA, it is not fit for analysis in R.

Thus, we refer Rob Hyndman's recommendation of using the fpp3 package's tsibble object. It allows us to manipulate time series with dplyr commands.

```{r}
## Convert to tsibble object 
US_QGDP$gdp <- as.numeric(US_QGDP$gdp)
US_QGDP$gdp <- format(round(US_QGDP$gdp,3), nsmall=3)


# Create quarterly data sequence
ti <- seq(as.Date("1947-01-01"), as.Date("2019-12-01"), by= "quarter")
length(ti) == length(US_QGDP$gdp) 

# Change to quarterly data
ti <- as.yearqtr(ti, format="%Y-%m-%d")
ti <- yearquarter(ti)
US_QGDP$yq <- ti

# Change character to numeric vector
US_QGDP$gdp <- as.numeric(US_QGDP$gdp)

# Swap column order, change to tsibble object
US_QGDP <- select(US_QGDP, yq, gdp) %>%
	as_tsibble(index = yq)

autoplot(US_QGDP)
```

## Q1

### i) detecting stationarity of series visually

Obviously, from the gdp plot, we see a clear trend, so the series is not stationary.

### ii) Take the Log difference

```{r}
plt.d.logUS_QGDP <- US_QGDP %>% transmute(
	'd.loggdp' = difference(log(gdp))
	) %>%
	
	ggplot(aes(x=yq, y= d.loggdp)) +
	geom_line(col='steelblue') +
	labs( y = "first difference of log gdp", title = "Growth rate of quarterly GDP")

plt.d.logUS_QGDP
```

### iii) estimate AR(1) model for GDP growth rate

```{r}

#Save differenced log series
d.logUS_QGDP <- US_QGDP %>% transmute(
	'd.loggdp' = difference(log(gdp))
  )
	
# Fit AR(1) model
fit <- d.logUS_QGDP %>% drop_na() %>%
	model(AR(d.loggdp ~ order(1), na.rm=T))

report(fit)
```

```{r}

#forecast 1 quarter ahead
fit %>% forecast(h = 1)

#Plot forecasted path and distribution along with original data
forecast.1 <- fit %>% forecast() %>% autoplot(d.logUS_QGDP) + 
	coord_x_date(xlim = c("2000-1-1" , "2021-1-1"))
forecast.1
```

### iv) Estimate AR(2) model for GDP Growth rate

```{r}
#Fit AR(2) model
fit2 <- d.logUS_QGDP %>% drop_na() %>%
	model(AR(d.loggdp ~ order(2), na.rm=T)
	)
	
report(fit2)

#Forecast 1 quarter ahead
fit2 %>% forecast(h = 1)

#Plot forecasted path along original data
forecast.2 <- fit2 %>% forecast() %>% autoplot(d.logUS_QGDP) +
coord_x_date(xlim = c("2000-1-1" , "2021-1-1"))

forecast.2

```
